Improving the Image Acquisition Rate \\ of an Atomic Force Microscope \\ Through Sub-sampling and Reconstruction

### 2019-06-24 latex/afm-cs-main.tex
Update on Overleaf.

\documentclass[twocolumn,oneside]{IEEEtran/IEEEtran}
% \documentclass[defaultstyle,11pt]{IEEETran}

\usepackage{showframe}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}		% to insert figures
\usepackage{multirow}
\usepackage{cite}
\usepackage{listings}
\usepackage{booktabs}
\usepackage[percent]{overpic}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage[inkscapelatex=false, inkscapepath=svgsubpath]{svg}
\usepackage{url}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{placeins}
\usepackage[disable]{todonotes}

@@

\begin{document}
\title{Improving the Image Acquisition Rate of an Atomic Force
  Microscope Through Sub-sampling and Reconstruction}

\author{Roger A. Braker, Yufan Luo, Lucy Y. Pao and Sean B. Andersson
  \thanks{The authors are with the Dept. of Electrical, Computer, and
    Energy Engineering at the University of Colorado, 425 UCB,
    Boulder, CO 80309, United States. Phone: +1 (303) 492-2360. Fax:
    +1 (303) 492-2758. R. A. Braker (corresponding author
    roger.braker@colorado.edu) is a graduate student and L.Y. Pao
    (pao@colorado.edu) is the Richard \& Joy Dorf Professor.}
  \thanks{This work was supported in part by the US National Science
    Foundation (NSF Grant CMMI-1234980) and Agilent Technologies,
    Inc.} }

@@

The Atomic Force Microscope (AFM) is a powerful instrument capable of
imaging sample surface topography, material characteristics, and other
surface properties, with nanometer scale precision. AFMs acquire
information about the sample by monitoring the deflection of the
cantilever caused by the tip-sample interaction force. In general,
% AFMs acquire information about the sample through a variety of
% imaging modes, all of which rely on the deflection of the cantilever
% caused by the tip-sample interaction force. While the details of
% each of the modes are quite different, in general,
feedback control is used to hold the deflection signal constant and
information about the property of interest is inferred from the
applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial
resolution, and ability to image in vacuum, air, and liquid, AFM is
widely used in a variety of disciplines, including physics, biology
and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}.

In addition to imaging static samples, AFM is used to study dynamics
in systems with nanometer-scale features
\cite{Yang:2017im,Shibata:2017da,Shibata:2015jd,Ando:2014ja}. The
image acquisition time of conventional AFMs, however, is typically on
the order of seconds to minutes, severely limiting the time-scales
that can be explored. Driven by the need for faster imaging, there
continue to be many active research efforts to overcome this
challenge. Since an AFM is fundamentally a mechanical microscope, many
of the approaches to high-speed AFM (HS-AFM) have focused on
modifications to the physical components. These have included the use
of small, fast cantilevers \cite{viani1999fast,
  braunsmann2010high,Adams:2016hg}, the development of faster
actuators \cite{Maroufi:2015gt,Yong:2012kd,Kenton:2012cm}, as well as
the application of advanced controllers
\cite{Rana:2018es,Yong:2015gr,butterworth2010adaptive,salapaka2002high}.
By combining different techniques, there are current generation,
high-end instruments that can image at rates on the order of 1-10
frames/sec \cite{Ando:2014ja}. However, the fastest rates are achieved
only over small scan ranges and there remain many systems of interest
whose dynamics are faster even than these instruments can achieve. In
addition, there is a large installed base of much slower AFMs that can
benefit from alternative approaches to improving imaging rate.

@@

Our previous work on the $\mu$-path pattern using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are:
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype, \cite{braker_hardware_2018} we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an affect which neither our prior experimental work nor the numerous simulation studies have accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to course raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a course raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{Luo_nano_2015,maxwell_acc_2014} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
\end{itemize}

@@

>>>>

\documentclass[journal]{IEEEtran}
% \documentclass[twocolumn,oneside]{IEEEtran/IEEEtran}
% \documentclass[defaultstyle,11pt]{IEEETran}

% \usepackage{showframe}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}		% to insert figures
\usepackage{multirow}
\usepackage{cite}
\usepackage{listings}
\usepackage{booktabs}
\usepackage[percent]{overpic}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage[inkscapelatex=false, inkscapepath=svgsubpath]{svg}
\usepackage{url}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{placeins}
\usepackage[disable]{todonotes}

@@

\begin{document}
\title{Improving the Image Acquisition Rate \\ of an Atomic Force Microscope \\
Through Sub-sampling and Reconstruction}

\author{Roger A. Braker, Yufan Luo, Lucy Y. Pao and Sean B. Andersson
  \thanks{The authors are with the Dept. of Electrical, Computer, and
    Energy Engineering at the University of Colorado, 425 UCB,
    Boulder, CO 80309, United States. Phone: +1 (303) 492-2360. Fax:
    +1 (303) 492-2758. R. A. Braker (corresponding author
    roger.braker@colorado.edu) is a graduate student and L.Y. Pao
    (pao@colorado.edu) is the Palmer  Professor.}
  \thanks{This work was supported in part by the US National Science
    Foundation (NSF Grant CMMI-1234980)  Agilent Technologies,
    Inc., and the Hanse-Wissenschaftskolleg Institute for Advanced Study in Delmenhorst, Germany.} }

@@

The Atomic Force Microscope (AFM) is a powerful instrument capable of
imaging sample surface topography, material characteristics, and other
surface properties, with nanometer scale precision. AFMs acquire
information about the sample by monitoring the deflection of the
cantilever caused by the tip-sample interaction force. In general,
% AFMs acquire information about the sample through a variety of
% imaging modes, all of which rely on the deflection of the cantilever
% caused by the tip-sample interaction force. While the details of
% each of the modes are quite different, in general,
feedback control is used to hold the deflection signal constant and
information about the property of interest is inferred from the
applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial
resolution, and ability to image in vacuum, air, and liquid, AFM is
widely used in a variety of disciplines, including physics, biology,
and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}.

In addition to imaging static samples, AFM is used to study dynamics
in systems with nanometer-scale features
\cite{Yang:2017im,Shibata:2017da,Shibata:2015jd,Ando:2014ja}. The
image acquisition time of conventional AFMs, however, is typically on
the order of seconds to minutes, severely limiting the time-scales
that can be explored. Driven by the need for faster imaging, there
continue to be many active research efforts to overcome this
challenge. Since an AFM is fundamentally a mechanical microscope, many
of the approaches to high-speed AFM (HS-AFM) have focused on
modifications to the physical components. These have included the use
of small, fast cantilevers \cite{Viani:1999tp,
  braunsmann2010high,Adams:2016hg}, the development of faster
actuators \cite{Maroufi:2015gt,Yong:2012kd,Kenton:2012cm}, as well as
the application of advanced controllers
\cite{Rana:2018es,Yong:2015gr,butterworth2010adaptive,salapaka2002high}.
By combining different techniques, there are current generation,
high-end instruments that can image at rates on the order of 1-10
frames/sec \cite{Ando:2014ja}. However, the fastest rates are achieved
only over small scan ranges and there remain many systems of interest
whose dynamics are faster even than these instruments can achieve. In
addition, there is a large installed base of much slower AFMs that can
benefit from alternative approaches to improving imaging rate.

@@

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are:
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an affect which neither our prior experimental work nor the numerous simulation studies have accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{Luo_nano_2015,maxwell_acc_2014} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
\end{itemize}

@@

\begin{IEEEbiography}{Roger A. Braker} is so happy that all this is almost done.

\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{yufan.JPG}}]{Yufan Luo}
received dual B.S. degrees in automation and in management science from Zhejiang University, Zhejiang, China, in 2011, an M.S. degree in industrial and operations engineering from the University of Michigan, Ann Arbor, MI, USA, in 2012, and a Ph.D. degree in systems engineering from Boston University in 2019. He is currently with Juniper Networks, Sunnyvale, CA.

\end{IEEEbiography}

\begin{IEEEbiography}{Lucy Y. Pao} is probably not so happy that Arnold's almost done.

\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{SeanAndersson_IEEE.jpg}}]{Sean B. Andersson}
received the B.S. degree in engineering and applied physics from Cornell University, Ithaca, NY, USA, in 1994, the M.S. degree in mechanical engineering from Stanford University, Stanford, CA, USA, in 1995, and the Ph.D. degree in electrical and computer engineering from the University of Maryland, College Park, MD, USA, in 2003. He has worked at AlliedSignal Aerospace and Aerovironment, Inc. and is currently an Associate Professor of mechanical engineering and of systems engineering with Boston University, Boston, MA, USA. His research interests include systems and control theory with applications in scanning probe microscopy, dynamics in molecular systems, and robotics. He received an NSF CAREER award in 2009, is a senior member of the IEEE, and was an associate editor for the IEEE Transactions on Automatic Control (2014-2018) and for the SIAM Journal on Control and Optimization (2013-2018).
\end{IEEEbiography}
